% File name: report/handbook.tex

\documentclass[a4paper,11pt]{article}  % Standard document class
\usepackage[english]{babel}            % Set document language
\usepackage{fullpage}                  % Set up page for small margins etc
\usepackage{listings}
\usepackage{graphicx}                  % For including images in document

\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

\usepackage{appendix}
\usepackage{color}
\usepackage{hyperref}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{parskip}

%\usepackage{placeins}                  % Allows use of \FloatBarrier
% to avoid images or tables
% moving into next section
%\usepackage{subfig}                    % For subfigures...

\usepackage{amsmath}                   % For improving maths/formula typesetting
%\usepackage{tabular}                  % Table changing package

%\usepackage{algpseudocode}             % For producing algorithms/flowcharts

\usepackage{listings}                  % For including source code in document
\lstset{
  basicstyle = \small
}

\setlength{\parindent}{0in}
%\setlength{\parskip}{0.3cm}

% Provide command for scientific notation
\providecommand{\e}[1]{\ensuremath{\times10^{#1}}}
\providecommand{\degrees}{\ensuremath{^{\circ}}}

% Define title here:
\title{Soldering Machine Handbook}
\author{James Glanville}
\date{15th May 2013}

\hypersetup{
    colorlinks,
    allcolors=blue,
    linktoc=all,
}

\begin{document}

% generate title
\maketitle

\tableofcontents

\newpage

\section{Introduction}

This handbook lists the steps required to operate the machine, and
produce populated PCBs. The operating system used in this handbook is
Ubuntu 13.10. The PCB CAD program used is Kicad.

\section{Prerequisites}
The following things are assumed:

\begin{itemize}
	\item	Computer with Windows/Linux. OS X may work, but is untested.
	\item	SD card (at least 128mb).
\end{itemize}

\section{Installing Software}
First install pcb2gcode as described in Appendix \ref{sec:buildpcb2gcode}.


\subsection{Project files}
Fetch the project's files from source control:
\begin{lstlisting}[frame=single,breaklines=true]
git clone https://github.com/JamesGlanville/solderingmachine.git
\end{lstlisting}

\subsection{Install pronterface (if PC control is required)}
\begin{lstlisting}[frame=single,breaklines=true]
git clone https://github.com/kliment/Printrun.git
\end{lstlisting}


\subsection{Preparing SD card}
Format the SD card to the FAT32 filesystem, ensuring the disk has an
MSDOS partition table and that the FAT32 partition is on the first
primary partition.

In Ubuntu, the following steps are necessary:

Install gparted if it is not already installed:

\begin{lstlisting}[frame=single,breaklines=true]
sudo apt-get install gparted
\end{lstlisting}

Launch gparted as root:
\begin{lstlisting}[frame=single,breaklines=true]
sudo gparted
\end{lstlisting}

Select the SD card using the drop down menu at the top right of the
screen.

Create the MSDOS partition table by clicking Device-\textgreater Create Partition Table.

Create the FAT32 filesystem by right-clicking on the empty disk, and selecting "New". 
Fill in the information in the dialogue box, and click "Apply changes". 

\subsection{Copying utility GCODE to SD card.}
Copy all .GCODE files from pcb2gcode-metric/GCODE/ to the root of the SD card.
These can be used to perform simple functions such as zeroing the axes.

\section{Creating GERBER files.}
The following steps should be followed to export GERBER files from Kicad:

\begin{enumerate}
	\item	Open the PCB layout in PCBNew.
	\item	Select \textbf{Tracks and Vias} from the \textbf{Dimensions} Menu.
	\item	Change the \textbf{Mask Clearance} to 0.0001.
	\item	Select \textbf{Plot} from the \textbf{File} menu.
	\item	Select the following layers: \textbf{Copper}, \textbf{Component},\textbf{SoldPCmp}, \textbf{EdgesPcb}
	\item	Click \textbf{Plot} and then \textbf{Generate drill file}.
\end{enumerate}

\section{Generating GCODE with pcb2gcode}
\subsection{Creating millproject file.}
An example millproject file is given in pcb2gcode-metric/millproject. This should be copied
into the directory containing the GERBER files, and modified if necessary (see included comments).

\subsection{Copy GCODE to SD Card}

\section{Operating the machine}
The following steps described the various stages involved in producing
a populated and soldered circuit board. Some steps (such as drilling)
may be omitted if not required.

\subsection{Attaching the PCB}
Attach copper-clad PCB to machine using double sided tape. It is important that the PCB is level,
so care must be taken ensure the tape is not bunched or folded. The PCB should be firmly
pressed to the bed to ensure a strong bond.

\subsection{Isolation Routing}
\begin{enumerate}

\item
Insert conical routing bit into the spindle. This is achieved by loosening
the idler section, and then loosening the two machine screws holding the bit in place.
The bit should be carefully removed by hand, and the conical routing bit 
pushed into its place. The two machine screws should be tightened, and the idler
tightened.

\item
The spindle tool should then be attached to the tool holder with the two
wing nuts.

\item
Switch on the machine.

\item
Insert the SD card containing all of the GCODE files and power on the 
machine.

\item
Run \textbf{zero.gcode} then \textbf{startspindle.gcode} then \textbf{mill.gcode}

\item
When routing is complete, run \textbf{stopspindle.gcode}

\end{enumerate}

\subsection{Drilling}
This section may be ignored if no holes need be drilled through the PCB.

\begin{enumerate}
\item
The drill bit should be inserted into the spindle tool as described in 
the "Isolation Routing" section above.

\item
Switch on the machine.

\item
Insert the SD card containing all of the GCODE files and power on the 
machine.

\item
Run \textbf{zero.GCODE} then \textbf{startspindle.GCODE} then \textbf{drill.GCODE}

\item
When drilling is complete, run \textbf{stopspindle.GCODE}

\end{enumerate}

\subsection{Paste Extrusion}

\begin{enumerate}

\item
Insert a syringe containing a sufficient volume of solder paste into
the paste extruder. Hand-tighten the extruder so that there is only a 
small amount of slack in the timing belt.

\item
Attach the paste extruder to the tool holder and hand-tighten the wing 
nuts.

\item
Switch on the machine.

\item
Insert the SD card containing all of the GCODE files and power on the 
machine.

\item
Run \textbf{zero.GCODE} then \textbf{prime.GCODE} then \textbf{paste.GCODE}

\end{enumerate}

\subsection{Placing components}
\begin{enumerate}
\item
Attach the vacuum needle tool to the machine and tighten with the wing nuts.

\item
Fill the part tray with the necessary SMD components.

\item
Switch on the machine.

\item
Insert the SD card containing all of the GCODE files and power on the 
machine.

\item
Run \textbf{zero.GCODE} then \textbf{place.GCODE}.
\end{enumerate}

\subsection{Reflowing the board}
\begin{enumerate}
\item
Attach the hot air tool to the machine and tighten with the wing nuts.

\item
Switch on the machine.

\item
Insert the SD card containing all of the GCODE files and power on the 
machine.

\item
Run \textbf{zero.GCODE} then \textbf{reflow.GCODE}.

\item
Run \textbf{cooldown.GCODE}

\end{enumerate}


\newpage
\appendix
\appendixpage
\addappheadtotoc

%\newpage
\section{Building pcb2gcode from source}
\label{sec:buildpcb2gcode}

The pcb2gcode source should first be obtained:
\begin{lstlisting}[frame=single,breaklines=true]
git clone git@github.com:JamesGlanville/pcb2gcode-metric.git
\end{lstlisting}

\subsection{Compilation in Linux}
For these instructions the use of Ubuntu/Debian is assumed, although other distributions should be similar.

To install prerequisites:
\begin{lstlisting}[frame=single,breaklines=true]
sudo apt-get install build-essential automake autoconf libtool libboost-all-dev libgtkmm-2.4-dev gerbv checkinstall
\end{lstlisting}

To compile pcb2gcode:
\begin{lstlisting}[frame=single,breaklines=true]
cd pcb2gcode-metric
./git-build.sh
\end{lstlisting}

(Optional) To install pcb2gcode:
\begin{lstlisting}[frame=single,breaklines=true]
sudo checkinstall
\end{lstlisting}

\subsection{Compilation in Windows}
Installing prerequisites is more complex in Windows:

\subsubsection{Visual Studio}
Install Visual Studio 2012 Professional from:
\begin{lstlisting}[frame=single,breaklines=true]
www.visualstudio.com
\end{lstlisting}

\subsubsection{GTK+ development files}
Download:
\begin{lstlisting}[frame=single,breaklines=true]
http://ftp.gnome.org/pub/gnome/binaries/win32/gtk+/2.24/gtk+-bundle_2.24.10-20120208_win32.zip
\end{lstlisting}
and extract to pcb2gcode-metric folder.

\subsubsection{Gerbv}
Download and install gerbvinst-2.6.1.exe from:
\begin{lstlisting}[frame=single,breaklines=true]
http://sourceforge.net/projects/gerbv/
\end{lstlisting}
%TODO: add instructions for generating .lib file

\subsubsection{Boost libraries}
%\begin{lstlisting}[frame=single,breaklines=true]
Download boost\_1\_55\_0.zip from:
\begin{lstlisting}[frame=single,breaklines=true]
http://sourceforge.net/projects/boost/
\end{lstlisting}

Unzip contents to c:\textbackslash boost155
open VS2012 x86 Native Tools Command Prompt, from:
\begin{lstlisting}[frame=single,breaklines=true]
 C:\ProgramData\Microsoft\Windows\Start Menu\Programs\Microsoft Visual Studio 2012\Visual Studio Tools
\end{lstlisting}

As administrator, run the following commands:
\begin{lstlisting}[frame=single,breaklines=true]
cd c:\boost155
bootstrap.bat
bjam.exe
\end{lstlisting}

\subsubsection{GTKMM}
Download and install gtkmm-win32-devel-2.22.0-2.exe from:
\begin{lstlisting}[frame=single,breaklines=true]
http://ftp.gnome.org/pub/GNOME/binaries/win32/gtkmm/2.22/
\end{lstlisting}

\subsubsection{Gerbv source}
Download gerbv-2.6.1.tar.gz from:
\begin{lstlisting}[frame=single,breaklines=true]
http://sourceforge.net/projects/gerbv/files/gerbv/gerbv-2.6.1/
\end{lstlisting}
and extract to c:\textbackslash gerbv-2.6.1

\subsubsection{Compilation}
The file pcb2gcode-metric.sln may now be opened in visual studio and the project can be compiled.

%hack appwindow.c
%create config.h
%create demo-config.h
%\end{verbatim}

\end{document}
